\chapter*{Acknowledgements}

First of all, I would like to thank my supervisor, Zoe~Kourtzi, for all the academic and mental support.
You managed to inspire me and gave me new insights in every conversation we had.
Thank you for turning me into a neuroscientist, and thank you for believing in me.
I also thank Andrew~Welchman for providing supervision and for his feedback.
Furthermore, I want to thank my advisor, Carola-Bibiane~Schönlieb, for her kindness and for being an inspiration, and together with Deborah~Vickers for providing valuable feedback on my first-year report.
I would like to thank Joana~Taylor~Tavares for creating an amazing and productive atmosphere in our lab, and for being patient with me.
I also thank Máté~Lengyel for inviting me to attend his journal club at the CBL, where I got a lot of new perspectives.
I warmly thank Mark~van~der~Wilk for his precise, insightful, and creative guidance on machine learning matters, not just on a technical level but also on a philosophical one.
I also want to thank Emmanuel~Stamatakis and Gustavo~Deco for taking the time to examine my thesis.

I am grateful to my close collaborators and friends Soroosh~Afyouni, Samuel~Bell, Joe~Ziminski, and Katharina~Zühlsdorff.
I have learned so much from working together with you.
You made me realize that science is a social endeavour, and I could not have done this without you.
I would also like to thank lab members Michael~Burkhart, Lorena~Santamaria~Covarrubias, Poly~Frangou, Máiréad~Healy, Vasilis~Karlaftis, Liz~Lee, Rui~Li, Elizabeth~Michael, Avraam~Papadopoulos, Reuben~Rideaux, Cecilia~Steinwurzel, Chie~Takahashi, Mengxin~Wang, and Elisa~Zamboni for welcoming me, for being my friend, and for making me feel at home in the lab.
I am grateful to Richard~Bethlehem and Varun~Warrier at the Department of Psychiatry for their incredibly professional and helpful support.

I thank Roland~Fleming and Kate~Storrs for welcoming me to Giessen and for providing a warm environment in the middle of winter.
I would also like to thank the entire Marie~Sk\l{}odowska-Curie DyViTo network for the cross-pollination and amazing conference in Cappadocia.
I express my gratitude to the European~Union's Horizon~2020 program and the respective taxpayers that indirectly funded this work.
Furthermore, I would like to thank the organizers of the machine learning summer school in T\"{u}bingen for giving me the opportunity to present an early iteration of this work, which allowed me to get valuable feedback.

I would also like to thank Queens' College, the Department of Psychology, the NHS, and the broader society in Cambridge for providing a community and being resilient in the face of the pandemic.

I also want to acknowledge all volunteers and everyone involved in data collection for the UK Biobank, the Human Connectome Project, and the Rockland sample, for putting in so much time and work.
Your contribution to science and progress is of immense value.

I thank my friends and family for being there for me.
Thank you, Fitz; you do not know me yet, but wanting to meet you has helped me a lot in making that final push.
Thank you Ink, for everything really.
And finally, a special thanks goes out to the large family of swans that visited me daily when I was writing up, for reminding me of the simple pleasures in life.
